\acknowledgement{
I would like to acknowledge all the people that made my doctorate possible.

My advisor, Joseph Devietti, for the opportunity to go to Penn and work with him. It has been a pleasure working and learning from your all these years. Joe is a great mentor, advisor, and incredible human being; he has been supportive and understanding all these years, specially when life got hard. I wish more of our projects would have panned out successfully. 

My best friend and closest collaborator, Kelly Shiptoski. Meeting you and working along side was the highlight of grad school. I hope we always continue working together.

All the Penn graduate students who made grad school fun! Our long, late, slightly drunk, conversations during TGIF. Special shout outs to Caleb Standford, Solomon Maina, Zach Schutzman, Alyssa Hwang, and, Irene Yoon. Thank you to the Penn PL group for allowing a systems person like me to join in.

Special mention to Irene Zhang, Microsoft researcher extraordinaire. She is an incredible mentor and one of the coolest people I know. 

My collaborators and mentors who gave me a chance and the opportunity to work with them: Ryan Newton, Alan Jeffrey, Leonid Ryzhyk, and Mihai Budiu.

Our research lab: Kelly Shiptoski (again!), Gautam Mohan, Bhavana Mehta, and Yuxuan Zhang.

Last but not least, my family and loved ones: My mother and father, Alma Leija Rea and Martin Barrera. My brother, Jose Navarro Leija. You guys supported me me all these years! And my lovely partner Dr. Hannah Joy Richter. Who's support was necessary to make it though the dissertation!
}